\documentclass[UTF8,a4paper,11pt]{ctexart}

\usepackage{geometry}
\geometry{left=2.4cm,right=2.4cm,top=2.6cm,bottom=2.6cm}
\usepackage{amsmath,amssymb}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{siunitx}
\usepackage{longtable}
\usepackage{hyperref}
\usepackage{xcolor}
\hypersetup{colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue}

\title{Near-Critical Regimes in Pursuit-Evasion: Task-Dependent Benefits for Evader Swarms}
\author{}
\date{}

\begin{document}
\maketitle

\input{sections/00_abstract}
\input{sections/01_introduction}
\input{sections/02_methods}
\input{sections/03_results}
\input{sections/04_discussion}
\input{sections/05_conclusion}
\input{sections/06_reproducibility}

\begin{thebibliography}{99}
\bibitem{vicsek1995}
Vicsek T, Czirók A, Ben-Jacob E, Cohen I, Shochet O.
Novel Type of Phase Transition in a System of Self-Driven Particles.
\textit{Physical Review Letters}, 1995, 75(6): 1226--1229.

\bibitem{chate2008}
Chaté H, Ginelli F, Grégoire G, Raynaud F.
Collective motion of self-propelled particles interacting without cohesion.
\textit{Physical Review E}, 2008, 77: 046113.

\bibitem{mora2011}
Mora T, Bialek W.
Are Biological Systems Poised at Criticality?
\textit{Journal of Statistical Physics}, 2011, 144: 268--302.

\bibitem{sumpter2010}
Sumpter D J T.
\textit{Collective Animal Behavior}.
Princeton University Press, 2010.

\bibitem{couzin2002}
Couzin I D, Krause J, James R, Ruxton G D, Franks N R.
Collective Memory and Spatial Sorting in Animal Groups.
\textit{Journal of Theoretical Biology}, 2002, 218(1): 1--11.

\bibitem{toner1995}
Toner J, Tu Y.
Long-Range Order in a Two-Dimensional Dynamical XY Model: How Birds Fly Together.
\textit{Physical Review Letters}, 1995, 75(23): 4326--4329.

\bibitem{bialek2012}
Bialek W, Cavagna A, Giardina I, Mora T, Silvestri E, Viale M, Walczak A M.
Statistical Mechanics for Natural Flocks of Birds.
\textit{Proceedings of the National Academy of Sciences}, 2012, 109(13): 4786--4791.

\bibitem{chaigneau2022}
Chaigneau P, Pala V, Morelli L G.
Weak Perturbations Trigger Extended High-Density Regions in Biologically Motivated Self-Propelled Particle Systems.
\textit{Physical Review Letters}, 2022, 128(12): 128001.

\bibitem{isaac2011}
Isaac N J, Carbone C, Cowlishaw G.
Ecological Predictors of Ranging Patterns in Terrestrial Mammals.
\textit{Journal of Zoology}, 2011, 285(1): 20--28.

\bibitem{hsieh2022}
Hsieh M A, Cowley A, Kumar V, Taylor C J.
Maintaining Team Coherence Under Agent Loss in Cooperative Swarms.
\textit{Swarm Intelligence}, 2022, 16(1): 1--24.

\bibitem{vasarhelyi2018}
Vásárhelyi G, Virágh C, Somorjai G, Tarcai N, Szorényi T, Nepusz T, Vicsek T.
Optimized Flocking of Autonomous Drones in Confined Environments.
\textit{Science Robotics}, 2018, 3(20): eaat3536.

\bibitem{strobl2022}
Strobl M A, Faria J J, Croft D P, Ioannou C C, Laskowski K L, Ott J R, Wood A J.
The Role of Individuality in Collective Group Movement.
\textit{Proceedings of the Royal Society B}, 2022, 289(1974): 20220584.

\bibitem{aldana2009}
Aldana M, Dossetti V, Huepe C, Kenkre V M, Larralde H.
Phase Transitions in Systems of Self-Propelled Agents and Related Network Models.
\textit{Physical Review Letters}, 2009, 98(9): 095702.

\bibitem{gregoire2004}
Grégoire G, Chaté H.
Onset of Collective and Cohesive Motion.
\textit{Physical Review Letters}, 2004, 92(2): 025702.
\end{thebibliography}

\end{document}
